On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence

### 2017-09-10 Thesis/Draft.Rmd
Add References

---
title: "Draft"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 11pt
urlcolor: blue
---

@@

@@

# Abbreviations

Kullback-Leibler divergence between distributions $\nu_i$ and $\nu_j$: $\KL(\nu_i||\nu_j)$.

Binary relative entropy: $d(x,y) := x \log(x/y) + (1-x) \log((1-x)/(1-y))$, where $d(0,0) = d(1,1) = 0$.

Kullback-Leibler divergence for two Bernoulli distributions parameterized by their means $\mu_i$ and $\mu_j$: $KL_{Ber}(\nu_i||\nu_j) = d(\mu_i, \mu_j)$.

Define exponential family of distributions

Define KL Divergence for this family

@@

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that as soon as possible more patients receive the better drug and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

@@

In this thesis, we consider the thresholding bandit problem. Instead of identifying the best $m$ out of $K$ arms, we aim to find all distributions out of $K$ that have a mean at least as large as a threshold fixed upfront. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above or below the threshold, and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

@@

To introduce the thresholding bandit problem, consider a game in which the learner is tasked to find the slot machines within a row of $K$ slot machines in a casino that give a mean payoff larger than a fixed threshold $\tau$, for example $\tau = 0$. The learner cannot pull all slot machines at the same time. Instead, he acquires samples sequentially from one slot machine at a time to learn about their payoff structure. The learner has a budget of $T>K$ slot machine pulls. After $t-1$ samples, the learner has observed $T_i(t-1)$ samples from slot machine $i$. The payoffs $X_{i,1}, ..., X_{i,T_i(t-1)}$ are drawn from the payoff distribution $\nu_i$ underlying the slot machine $i$. At time $t$, the learner considers the payoffs he observed so far in order to choose the next slot machine he would like to pull. Her decision is lead by the goal to classify the slot machines with distribution $\nu_i$ and mean payoff $\mu_i = \mathbb{E}_{X \sim \nu_i}[X]$ after $T$ samples as being above or below the threshold with error probability as small as possible.

The thresholding bandit problem can thus be framed in the literature on sequential sampling strategies. Even though its name is reminiscient of the multi-armed bandit setting, the thresholding bandit problem is more akin to best arm identification and top-$m$ identification problems due to its focus on exploration: While the classical multi-armed bandit problems are concerned with cumulative regret which increases *during* sampling, best arm identification and thresholding bandits collect loss only once *after* the sampling procedure (compare Garivier and Kaufmann, 2016; Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013). As we will see in \autoref{chap:KullbackLeiblerPureExploration} there arises an implicit threshold between the means of the $m$-th best arm and the $m+1$-th best arm in the top-$m$ identification problem (Kaufmann et al., 2016; Kaufmann and Kalyanakrishnan, 2013). Because of this, the two settings are very related.

## Setup \label{sec:Setup}

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, she would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which she would like to observe the reward $X_{I_t,t}$. Often, adaptive sampling strategies makes the decision of which arm to pull by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then is $B_i(t)$ a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

@@

### Proof of Theorem \ref{theorem:LocatelliTheorem4}

@@

Given that we do not follow a uniform sampling strategy, we know that at time $T$, some arms will have been pulled more frequently than others. In particular, some arms will have been pulled more than $T/K$ times. How often an arm is pulled can be related to the individual arm's contribution to the overall complexity measure. Let $h_i$ be a measure of arm $i$'s individual problem complexity and write for the overall problem complexity $H = \sum_{i = 1}^{K} h_i$. Then arm $i$ contributes a share of $h_i/H$ to the overall problem complexity. We can now show that there exists an arm that has been pulled at least as often as the share of his complexity contribution.

@@

### Check Separation Condition

@@

If we write the lower bound as $LB = 12(\log(T)+1)K)\exp(-3T/H)\exp(-4) = C_{LB}(\log(T)+1)K\exp(-T/H)$ and the upper bound as
$UB = (\log(T)+1)K \exp(2) \exp(-\frac{1}{64R^2}\frac{T}{H}) = C_{UB}(\log(T)+1)K\exp(-T/H)$, we see that $LB$ and $UB$ match up to multiplicative constants and the $(\log(T)+1)K$ term which is negligiable in comparison with the exponent for large $T$ as shown in Locatelli et al. (2016), which makes the APT algorithm order optimal.

@@

- Mention Optimality (Upper Bound matches lower bound)
- Anytime

One of the APT algorithm's main features is its simplicity. The sampling rule is very straightforward, and, as highlighted in Locatelli et al. (2016), it does not require any knowledge about the type of sub-Gaussianity $R$, or the complexity $H$ upfront. This makes the algorithm immediately applicable in practice while it still enjoys its upper bound on the expected regret. This is in particular worth highlighting when comparing APT to algorithms used in the best arm identification or Top-$m$ setting. As alluded to in \autoref{sec:TheAPTAlgorithm}, many algorithms that fix the sampling budget upfront (i.e., fixed budget algorithms as UCB-E or KL-LUCB-E) require knowledge about the corresponding complexity upfront (Audibert et al., 2010; Kaufmann and Kalyanakrishnan, 2013). The problem however being, that to know the complexity, the experimenter has to know all means up to a permutation of the arms. This, of course, is not given in practice. Thus alternative algorithms try to estimate the complexity online which can be difficult as especially in the beginning plug-in estimates based on the estimated means may lead to wrong estimates of the complexity, which in turn leads to poor budget allocation based on mistaken beliefs (compare Garivier and Kaufmann, 2016). On the other hand, the sequential sampling strategy has to rely on some estimate. While the APT algorithm does not explicitly estimate the complexity, the gaps that serve as index are essentially the same. Faced not with sub-Gaussian but more heavy-tailed distributions, these estimates might well lead to poor performance.

Here, the APT algorithm shows room for improvement, as also mentioned in Locatelli et al. (2016). The problem is that the assumptions for the bounds no longer hold as arms are less sub-Gaussian; of course the complexity $H$ and the term $\hat{\Delta}_i(t)$ in the algorithm's index are motivated by sub-Gaussian distributions. The more one moves away from those, the less appropriate these quantities become. In particular, the empirical mean will be less (or not at all) concentrated around the true mean and more impacted by outliers which are more likely under non sub-Gaussian distributions. Locatelli et al. (2016) thus refers to Bubeck et al. (2013a) which considers solutions for multi-armed bandits that are based on heavy-tail distributions. By using more robust estimators of the mean, one could equip the APT algorithm to deal with such distributions.

Both Mukherjee et al. (2017) as well as Zhong et al. (2017) use estimates of the variance of the arms for their algorithms for the thresholding bandit problem and try to thereby improve upon the performance of the APT algorithm. While their simulations show promising results, their bounds rely on the assumption of bounded reward distributions. Thus their algorithms are not necessarily able to improve upon the assumption of sub-Gaussian distributions made by Locatelli et al. (2016). We will continue to discuss their results in \autoref{chap:VarianceBasedAlgorithms}.

@@

$$
\KL_{N}(\nu_i || \nu_j) = 2(\mu_i - \mu_j)^2.
$$

@@

$$
\KL_{Ber}(\nu_i||\nu_j) = \mu_i \log(\frac{\mu_i}{\tau}) + (1-\mu_i) \log(\frac{1-\mu_i}{1-\tau})
$$

@@

$$
\KL_{Ber}(\nu_i||\nu_j) \geq 2(\mu_i - \tau)^2.
$$

This is of course equal to the Gaussian KL-divergence described before. This approximation, however, suffers for heavily for very small values of $\mu_i$ and $\tau$. In particular, the Bernoulli KL-divergence becomes asymmetric in its first argument as $\tau \neq 0.5$. Thus, in contrast to the complexity measure $H$ based on the Gaussian KL-divergence, a complexity measure based on the Bernoulli KL-divergence would evaluate two arms as differently difficult if their absolute distance to the threshold is the same, but they lie on different sides of the threshold.

Thus similarly to how UCB algorithms in the cumulative regret setting have been extended to KL-UCB algorithms (compare Cappé et al., 2013), also the APT algorithm offers room for improvement. As a more explicit example of such adaptation, consider the case of Top-$m$ identification described in Kaufmann et al. (2016). The authors develop a lower bound based on KL-divergence, and note that "This information-theoretic lower bound permits to interpret the quantity $H(\nu)$ [...] as a subgaussian approximation". Given that Bernoulli distributions with very small parameter values arise naturally in many applications, an improvement in this regard would be valuable.

In what follows, we first examine in \autoref{chap:KullbackLeiblerPureExploration} how information theoretic quantities are employed in the best arm identification and Top-$m$ problem as detailed in Kaufmann and Garivier (2016) and Kaufmann et al. (2016). Before introducing a new algorithm for the thresholding bandit problem based on Kullback-Leibler divergences in \autoref{sec:LikelihoodRatioForExponentialFamily}, we first describe the variance-based approaches in \autoref{sec:VarianceBasedAlgorithms}.

@@

ggplot(kldf_total, aes(x = mu, y = KL, 
                      group = interaction(KL_Type, tau))) +
  facet_grid(level~., scales = "free") +
  geom_line(aes(color = tau, linetype = KL_Type)) +
  labs(x = "First argument of KL-divergence",
       y = "KL-divergence",
       title = "Comparison of Bernoulli KL-divergence and 
Gaussian Approximation")  +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_discrete(name = "Second argument of KL-divergence") +
  scale_linetype_discrete(name = "Type of KL-divergence") +
  guides(color = guide_legend(title.position = "top"),
         linetype = guide_legend(title.position = "top"))
```

@@

The previous chapter introduces the thresholding bandit problem as described in Locatelli et al. (2016), and with the APT strategy a first algorithm that is simple, yet order-optimal for large budgets. However, we end with the conclusion that a distance measure between the distributions and the threshold based on the Kullback-Leibler divergence might be more useful than the simple gap $\Delta^2_i = (\mu_i - \tau_i)^2$ employed so far. Indeed, citing Kaufmann et al. (2016), we note that the complexity $H$ used so far can be considered a sub-Gaussian approximation of a better complexity measure based on information theoretic measures like the Kullback-Leibler divergence.

In order to give more intuition to why that is the case, and in what ways information theoretic measures can be introduced to measure the complexity of bandit problems, we now turn to Kaufmann et al. (2016) and Garivier and Kaufmann (2016). The latter describes a strategy based on Kullback-Leibler divergences for the best-arm identification problem, while the first derives complexity measures based on the divergence. Ideally, their work offers methods that can inform our work of extending the thresholding bandit problem. In what follows, we adopt the notation of Kaufmann et al. (2016) to ease a comparison with their arguments.

@@

Instead of the thresholding bandit problem, Kaufmann et al. (2016) consider the problem of identifying the best $m$ out of $K$ arms, that is, the top-$m$ problem. They describe their bandit model $\nu$ by $K$ arms characterized by probability distributions $\nu_a$, $a \in \{1,...,K\}$, with mean $\mu_a$. A learner picks at any time step $t=1,2,...,$ an arm $A_t \in \{1,...,K\}$ for which he observes the independent reward $Z_{t} \sim \nu_{A_t}$. The learner aims to find the $m$ out of $K$ arms with the largest expected reward. This set of arm indexes is denoted by $\mathcal{S}_m^*$. Kaufmann et al. write the means $(\mu_{1}, ..., \mu_{K})$ sorted in decreasing order as the K-tuple $(\mu_{[1]}, ..., \mu_{[K]})$. Based on this, they require that $\mu_{[m]} > \mu_{[m+1]}$ such that the solution $\mathcal{S}_m^*$ is unique. The class of bandit problems $\nu$ for which the latter condition holds is called $\mathcal{M}_m$ by Kaufmann et al. At the end of the sampling procedure, after $T$ samples (denoted by $\tau$ in Kaufmann et al.) the learner is required to return a classification $\hat{\mathcal{S}}_m$. This classification is then evaluated by the loss function $\mathcal{L}(T) = \mathbb{I}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m)$. This binary loss is similar to the loss in the thresholding bandit. We can now define a strategy of the learner as $\mathcal{A} = ((A_t), T, \hat{\mathcal{S}}_m)$: The strategy defines how an arm is chosen, how many samples are observed, and how the arms are eventually classified.

When introducing the thresholding bandit setting, we did not mention the fact that the budget $T$ does not need to be fixed upfront. Indeed, the setting in which the budget is *not* fixed upfront might be more popular among the best arm identification and top-$m$ literature. This setting is called *fixed confidence*, while the setting we considered so far is called *fixed budget*. In the fixed budget setting, we run algorithms for $T$ samples and then compare them based on their expected loss, $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m) =: p_t(\nu)$ as we have seen in $\autoref{sec:Setup}$. Alternatively, consider a $\delta$-PAC strategy $\mathcal{A}(\delta)$for which it holds for every problem $\nu \in \mathcal{M}_m$ that the classification holds at least with confidence level $1-\delta$, that is, $\mathbb{P}(\mathcal{S}^*_m = \hat{\mathcal{S}}_m) \geq 1-\delta$. The strategy has to achieve this level of confidence in its prediction before it is allowed to stop sampling. This is the fixed confidence case, in which the budget $T$ is no longer fixed upfront, but turns into a random variable. Algorithms are then compared by their expected number of draws $\mathbb{E}_{\nu}[T_{\delta}]$ and we prefer those which achieve the confidence level $1-\delta$ with a smaller expected budget.

In the fixed confidence setting, a family of strategies $A = (\mathcal{A}(\delta))_{\delta \in (0,1)}$ is called PAC if for every $\delta$, $\mathcal{A}(\delta)$ is $\delta$-PAC. On the other hand, in the fixed budget setting, a family of strategies $A = (\mathcal{A}(t))_{t \in \mathbb{N}}$ is called consistent if $p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$ for every problem $\nu \in \mathcal{M}_m$.

>>>>

---
title: "On Pure Exploration, Thresholding Bandits, and Kullback-Leibler Divergence"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{corollary}{Corollary}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   - \newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 11pt
urlcolor: blue
---

@@

\vspace*{10\baselineskip}

@@

@@

This makes sequential sampling especially relevant in situations where there is an inherent cost to obtaining samples, or where only a limited amount of samples can be collected. The classical example is that of clinical trials in which a researcher tests two drugs on patients to observe which one performs better. Not only is it difficult to gather enough patients, but one would also like to adapt the sampling such that as soon as possible more patients receive the better drug and are cured. As noted in Cappé et al. (2013), this example dates back to Thompson (1933, 1935). Such problems are known as multi-armed bandit problems in which the sampling is adapted so as to minimize the cumulative regret: In the case of the drug tests, one would like to sample the two drugs in a way to find the best drug as soon as possible and then to test this drug more and more often and more often than the worse drug. This would lead to cumulatively more healthy patients than testing both drugs uniformly; it would minimize the regret from treating patients with an ineffective drug.

@@

In this thesis, we consider the thresholding bandit problem. While the name is reminiscient of the cumulative regret multi-armed bandit, the problem is more similar to the top-$m$ problem. But instead of identifying the best $m$ out of $K$ arms, one aims to find all distributions out of $K$ that have a mean at least as large as a fixed threshold. Given that an implicit threshold arises in the top-$m$ problem between the $m$-th best and the $(m+1)$-th best arm (Kaufmann et al., 2016), the two problems share many characteristics. We discuss this in more detail in \autoref{chap:KullbackLeiblerPureExploration}. In the form described in Locatelli et al. (2016), algorithms in the thresholding bandit problem are evaluated against a binary loss function penalizing whenever an arm has not been classified correctly as being above or below the threshold, and thus focus on exploration similar to algorithms employed for best arm identification. Indeed, Locatelli et al. (2016) show how their thresholding algorithm can be used for best arm identification when the mean of the best arm is known upfront.

@@

## An Example

To introduce the thresholding bandit problem, consider a common situation among internet-based companies. These companies often collect user feedback to offer reviews and ratings on websites or in apps, or collect it to improve their services. Netflix lets users rate movies with "Thumbs up" and "Thumbs down", apps like Clue ask users a question per day to learn to predict the user's health and mood. An online shop might give customers a one-question survey after they make a purchase. Finally, Foursquare and Yelp ask users to rate coffee shops, bars, and restaurants.

Consider the problem the latter companies face. Whenever a user reports that she visited a restaurant, the app would like to ask questions about her experience. To give a simple example, assume users are not willing to answer more than one question per restaurant visit, and the app would like to find out two things: Does the restaurant have good service? Is the restaurant trendy?

If the feedback that users can give is restricted to binary "Yes" or "No" answers, then we can model the feedback as Bernoulli random variables

$$
X_{\text{service}} \sim \text{Ber}(\mu_{\text{service}}) \quad \text{and} \quad X_{\text{trendy}} \sim \text{Ber}(\mu_{\text{trendy}}).
$$

The app will then display a "Good service" or "This place is trendy" description for the restaurant if $\mu_{\text{service}}$ and $\mu_{\text{trendy}}$ are estimated to be above a certain threshold, for example $\tau = 0.5$. That is, more than half of the people would describe the place as having good service, or as trendy.

To make the app as useful as possible to other users, this information should be shown as quickly as possible. But even though the app has thousands of users, not that many people visit a specific restaurant, and even less people report this back to the app. Thus the company only has a limited budget of $T$ potential users from which to collect samples $X_{\text{service}}$ and $X_{\text{trendy}}$ from. How can the company ask these $T$ questions to be as confident as possible in its classification, that is, in the information it displays?

A straightforward way of sampling users would be to ask the two questions in a round-robin fashion: Ask for "Service", then for "Trendy", then for "Service", ... Given 100 users, 50 users give feedback on the service level, 50 users report on the trendiness level. This however is unlikely to be optimal with respect to the confidence after 100 samples. Given that the company asks the users sequentially, there is the option to adapt the sampling strategy over time: If the first ten users report $\sum_{s=1}^{10}X_{\text{service},s}=10$, while $\sum_{s=1}^{10}X_{\text{trendy},s}=5$, then the company might already be quite confident that $\mu_{\text{service}} \geq \tau = 0.5$, while this is less clear for $\mu_{\text{trendy}}$. Instead of continuously sampling both questions, it might be more useful to collect more feedback on the trendiness level so as to figure out its position relative to the threshold. After all, the company would like to use the 100 samples to maximize its confidence in both displayed tips.

Finding the optimal sequential sampling strategy to classify distributions against a threshold with as much confidence as possible after a limited budget of samples is the problem described above, and the problem described by the thresholding bandit problem which we discuss for the remainder of this thesis. In the following section, we define the general setting of the thresholding bandit problem.

## General Setting \label{sec:Setup}

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call the distributions arms, and observations from the arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, she would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from the different arms, where $X_{i} \sim \nu_i$. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which she would like to observe the reward $X_{I_t,t}$. Often, adaptive sampling strategies makes the decision of which arm to pull by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then is $B_i(t)$ a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indexes can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

@@

### Proof of Theorem \ref{theorem:LocatelliTheorem4} \label{sec:UpperBoundProof}

@@

Given that we do not follow a uniform sampling strategy, we know that at time $T$, some arms will have been pulled more frequently than others. In particular, some arms will have been pulled more than $T/K$ times. How often an arm is pulled can be related to the individual arm's contribution to the overall complexity measure. Let $h_i$ be a measure of arm $i$'s individual problem complexity and write for the overall problem complexity $H = \sum_{i = 1}^{K} h_i$. Then arm $i$ contributes a share of $h_i/H$ to the overall problem complexity. We can now show that there exists an arm that has been pulled at least as often as the share of its complexity contribution.

@@

### Check Separation Condition \label{sec:CheckSeparationCondition}

@@

If we write the lower bound as $LB = 12(\log(T)+1)K)\exp(-3T/H)\exp(-4) = C_{LB}(\log(T)+1)K\exp(-T/H)$ and the upper bound as
$UB = (\log(T)+1)K \exp(2) \exp(-\frac{1}{64R^2}\frac{T}{H}) = C_{UB}(\log(T)+1)K\exp(-T/H)$, we see that $LB$ and $UB$ match up to multiplicative constants and the $(\log(T)+1)K$ term which is negligiable in comparison with the exponent for large $T$ as shown in Locatelli et al. (2016). This makes the APT algorithm order optimal.

@@

One of the APT algorithm's main features is its simplicity. The sampling rule is straightforward, and, as highlighted in Locatelli et al. (2016), it does not require any knowledge about the type of sub-Gaussianity $R$, or the complexity $H$ upfront. This makes the algorithm immediately applicable in practice while it still enjoys its upper bound on the expected regret. This is in particular worth highlighting when comparing APT to algorithms used in the best arm identification or top-$m$ setting. As alluded to in \autoref{sec:TheAPTAlgorithm}, many algorithms that fix the sampling budget upfront (i.e., fixed budget algorithms as UCB-E or KL-LUCB-E) require knowledge about the complexity upfront (Audibert et al., 2010; Kaufmann and Kalyanakrishnan, 2013). The problem however being that to know the complexity, the experimenter has to know all means up to a permutation of the arms. This, of course, is not given in practice. Thus alternative algorithms try to estimate the complexity online which can be difficult as especially in the beginning plug-in estimates based on the estimated means may lead to wrong estimates of the complexity, which in turn leads to poor budget allocation based on mistaken beliefs (compare Garivier and Kaufmann, 2016). On the other hand, the sequential sampling strategy has to rely on some estimate. While the APT algorithm does not explicitly estimate the complexity, the gaps that serve as index are essentially the same. Empirically, we see in \autoref{chap:Experiments} that the problem described by Garivier and Kaufmann (2016) does also occur for the APT strategy. Faced not with sub-Gaussian but with more heavy-tailed distributions, this sensitivity is likely to increase.

Furthermore, as mentioned in Locatelli et al. (2016), the assumptions for the bounds no longer hold as arms become less sub-Gaussian; the complexity $H$ and the term $\hat{\Delta}_i(t)$ in the algorithm's index are both motivated by sub-Gaussian distributions. The more a problem deviates from those, the less appropriate the quantities become. In particular, the empirical mean will be less (or not at all) concentrated around the true mean and more impacted by outliers which are more likely under non sub-Gaussian distributions. Locatelli et al. (2016) thus refers to Bubeck et al. (2013a) which consider solutions for multi-armed bandits that are based on heavy-tail distributions. By using more robust estimators of the mean, one could equip the APT algorithm to deal with such distributions.

Both Mukherjee et al. (2017) as well as Zhong et al. (2017) use estimates of the variance of the arms for their algorithms for the thresholding bandit problem and try to thereby improve upon the performance of the APT algorithm. While their simulations show promising results, their bounds rely on the assumption of bounded reward distributions. Thus their algorithms are not necessarily able to improve upon the assumption of sub-Gaussian distributions made by Locatelli et al. (2016). We will continue to discuss their results in \autoref{sec:VarianceBasedAlgorithms}.

@@

$$
\KL_{\mathcal{N}}(\nu_i || \nu_j) = 2(\mu_i - \mu_j)^2.
$$

@@

$$
\KL_{\text{Ber}}(\nu_i||\nu_j) = \mu_i \log(\frac{\mu_i}{\tau}) + (1-\mu_i) \log(\frac{1-\mu_i}{1-\tau})
$$

@@

$$
\KL_{\text{Ber}}(\nu_i||\nu_j) \geq 2(\mu_i - \tau)^2.
$$

This is equal to the Gaussian KL-divergence described before. The approximation, however, suffers heavily for very small values of $\mu_i$ and $\tau$. In particular, the Bernoulli KL-divergence becomes asymmetric in its first argument when $\tau \neq 0.5$. Thus, in contrast to the complexity measure $H$ based on the Gaussian KL-divergence, a complexity measure based on the Bernoulli KL-divergence would evaluate two arms as differently difficult if their absolute distance to the threshold is the same, but they lie on different sides of the threshold.

Thus similarly to how UCB algorithms in the cumulative regret setting have been extended to KL-UCB algorithms (Cappé et al., 2013), the APT algorithm offers room for improvement, too. As a more explicit example of such adaptation, consider the case of top-$m$ identification described in Kaufmann et al. (2016). The authors develop a lower bound based on KL-divergence, and note that "This information-theoretic lower bound permits to interpret the quantity $H(\nu)$ [...] as a subgaussian approximation". Given that Bernoulli distributions with very small or very large parameter values arise naturally in many applications, an improvement in this regard would be valuable.

In what follows, we first examine in \autoref{chap:KullbackLeiblerPureExploration} how information theoretic quantities are employed in the best arm identification and top-$m$ problem as detailed in Kaufmann and Garivier (2016) and Kaufmann et al. (2016). Before introducing a new algorithm for the thresholding bandit problem based on Kullback-Leibler divergences in \autoref{sec:LikelihoodRatioForExponentialFamily}, we first describe the variance-based approaches in \autoref{sec:VarianceBasedAlgorithms}.

@@

ggplot(kldf_total, aes(x = mu, y = KL, 
                      group = interaction(KL_Type, tau))) +
  facet_grid(level~., scales = "free") +
  geom_line(aes(color = tau, linetype = KL_Type)) +
  labs(x = "First argument of KL-divergence",
       y = "KL-divergence"#,
       #title = "Comparison of Bernoulli KL-divergence and Gaussian Approximation"
       )  +
  theme_bw() +
  #theme(legend.position = "bottom") +
  scale_color_discrete(name = "Second argument \nof KL-divergence") +
  scale_linetype_discrete(name = "Type of \nKL-divergence") +
  guides(color = guide_legend(title.position = "top"),
         linetype = guide_legend(title.position = "top"))
```

@@

The previous chapter introduces the thresholding bandit problem as described in Locatelli et al. (2016), and the APT strategy as a first algorithm that is simple, yet order-optimal for large budgets. However, we end with the conclusion that a KL-divergence based measure of the distance between the distributions and the threshold might be more useful than the simple gap $\Delta^2_i = (\mu_i - \tau_i)^2$ employed so far. Indeed, citing Kaufmann et al. (2016), we note that the complexity $H$ -- and thus the gaps -- used so far can be considered a sub-Gaussian approximation of a better complexity measure based on information theoretic measures like the Kullback-Leibler divergence.

In order to give more intuition to why that is the case, and in which ways information theoretic measures can be introduced to measure the complexity of bandit problems, we now turn to Kaufmann et al. (2016) and Garivier and Kaufmann (2016). The latter describes a strategy based on Kullback-Leibler divergences for the best-arm identification problem, while the first derives complexity measures based on the divergence. Their work offers methods that can inform our goal of extending the thresholding bandit problem. In what follows, we adopt the notation of Kaufmann et al. (2016) to ease a comparison with their arguments.

@@

Instead of the thresholding bandit problem, Kaufmann et al. (2016) consider the problem of identifying the best $m$ out of $K$ arms, that is, the top-$m$ problem. They describe their bandit model $\nu = (\nu_1,...,\nu_K)$ by $K$ arms characterized by probability distributions $\nu_i$, $i \in \mathbb{A} = \{1,...,K\}$, with mean $\mu_i$. A learner picks at any time step $t=1,2,...,$ an arm $I_t \in \{1,...,K\}$ for which she observes the independent reward $X_{t} \sim \nu_{I_t}$. The learner aims to find the $m$ out of $K$ arms with the largest expected reward. This set of arm indexes is denoted by $\mathcal{S}_m^*$. Kaufmann et al. write the means $(\mu_{1}, ..., \mu_{K})$ sorted in decreasing order as the K-tuple $(\mu_{[1]}, ..., \mu_{[K]})$. Based on this, they require that $\mu_{[m]} > \mu_{[m+1]}$ such that the solution $\mathcal{S}_m^*$ is unique. The class of bandit problems $\nu$ for which the latter condition holds is called $\mathcal{M}_m$ by Kaufmann et al. At the end of the sampling procedure, after $T$ samples, the learner is required to return a classification $\hat{\mathcal{S}}_m$. This classification is then evaluated by the loss function $\mathcal{L}(T) = \mathbb{I}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m)$. This binary loss is similar to the loss in the thresholding bandit. We can now define a strategy of the learner as $\mathcal{S} = ((I_t), T, \hat{\mathcal{S}}_m)$: The strategy defines how an arm is chosen, how many samples are observed, and how the arms are eventually classified.

When introducing the thresholding bandit setting, we did not mention the fact that there exists an alternative setting in which the budget $T$ is not fixed upfront. This setting is called *fixed confidence*, while the case we considered so far is called *fixed budget*. In the fixed budget setting, we run algorithms for $T$ samples and then compare them based on their expected loss, $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}(\mathcal{S}^*_m \neq \hat{\mathcal{S}}_m) =: p_t(\nu)$ as we have seen in $\autoref{sec:Setup}$. Alternatively, consider a $\delta$-PAC strategy $\mathcal{S}(\delta)$for which it holds for every problem $\nu \in \mathcal{M}_m$ that the classification holds at least with confidence level $1-\delta$, that is, $\mathbb{P}(\mathcal{S}^*_m = \hat{\mathcal{S}}_m) \geq 1-\delta$. The strategy has to achieve this level of confidence in its prediction before it is allowed to stop sampling. This is the fixed confidence case, in which the budget $T=T_{\delta}$ is no longer fixed upfront, but becomes a random variable. Algorithms are then compared by their expected number of draws $\mathbb{E}_{\nu}[T_{\delta}]$ and we prefer those which achieve the confidence level $1-\delta$ with a smaller expected budget.

In the fixed confidence setting, a family of strategies $S = (\mathcal{S}(\delta))_{\delta \in (0,1)}$ is called PAC if for every $\delta$, $\mathcal{S}(\delta)$ is $\delta$-PAC. On the other hand, in the fixed budget setting, a family of strategies $S = (\mathcal{S}(t))_{t \in \mathbb{N}}$ is called consistent if $p_t(\nu) \rightarrow 0$ as $t \rightarrow \infty$ for every problem $\nu \in \mathcal{M}_m$.

Given this setup, we can now follow the arguments of Kaufmann et al. (2016) to derive complexity measures for the top-$m$ problem. As alluded to earlier, these complexity measures will rely on the KL-divergence in contrast to the complexity $H$ used by Locatelli et al. (2016). When discussing $H$, its motivation is that it represents the difficulty of the problem by measuring the distance between arms and the threshold. Instead, Kaufmann et al. approach complexity by how many samples (what error probability) one will need in the fixed confidence (fixed budget) setting. In order to arrive at their new measure of complexity, it makes sense to first follow their arguments to a general lower bound on the sample complexity $\mathbb{E}_{\nu}[T]$ which is given in their Theorem 4. Here, $\mathcal{P}$ refers to a set of probability measures for which Kaufmann et al. (2016) assume certain continuity conditions in their Assumption 3, and $\nu_i \in \mathcal{P}$. In the following, $d(x,1-x) = x\log(\frac{x}{1-x}) + (1-x) \log(\frac{1-x}{x})$ is the binary KL-divergence.
